Replacing handwritten signatures with open electronic signature software

### 2020-05-18 paper/main.tex
Fix additional proofreading issues

\abstract{
With the recent changes in the legal status of electronic signatures in the world, there is an opportunity to expand the use of electronic signatures as an alternative to handwritten signatures outside the usual areas.
In this work, we devote our attention to the exploration of the legal preconditions relevant for the EU, technological preconditions, review of the existing software, and a proposal of our own software platform.
The result is an open software platform for operating systems Windows, macOS, and Linux.
Using this application, it is possible, following the eIDAS Regulation, to sign and verify the signature of any document with the legal effect of an attested signature.
Our solution supports ID cards of the Slovak Republic with the eID chip without any modifications or settings, which enables use outside the web applications of the state.
The application is not tied only to the Slovak eID but is ready for further use with various devices in other countries and as a platform for other types of signatures.
There is also a website that, in addition to the possibility of downloading software and brief documentation, provides general information about electronic signatures.
The emphasis is on the simple extensibility of the program via language translations, documentation, functionality, and general information.
The thesis can be further extended with novel ways of signing or by improving flexibility, which is not completely up to par with compared software.
}

@@

While, specifically in the European Union, electronic signatures seem to be used relatively extensively in communication with the government, they do not seem to have been widely adopted in exchange between individuals or smaller organizations, which still seem to prefer handwritten signatures.
In many cases, however, they may be impractical for electronic communication, due to the need of a person's physical presence or the printing and sending of documents.

Our goal is to explore an accessible way of using electronic signatures.
In the beginning, we go over the preconditions that make it a viable alternative to handwritten signatures.
We cover the legal preconditions relevant to the European Union, and Slovakia in particular.
Also, we explain technical preconditions, focusing on the cryptographic concepts like asymmetric cryptography, timestamping, hashing, and standards that comply with the laws of the European Union.
Secondly, we do a review of the existing software, where we briefly cover several applications that are meant for basic use.
We compare them using several criteria to find out the strengths and weaknesses of the current implementations and identify a potential gap.
Finally, we devote our attention to our own software's implementation.
We explain the proposed desktop software and go into details on software engineering aspects.
Specifically, we cover usability, localization, modularity, architecture, testing, and automation.
In the end, we compare it to the existing software and identify possible future improvements.
Another result we focus on is a supporting website and how it helps with the use of the software, all in a way that is open for extension by the community.

@@

We did a review of the existing software, focusing on the licensing, cross-platform compatibility, localization, file support, ease of setup, flexibility, verification ability, and extensibility.
We identified the positive and negative sides of the current implementations and found a gap in the software for signing using qualified electronic signatures.

>>>>

\abstract{
With the recent changes in the legal status of electronic signatures in the world, there is an opportunity to expand the use of electronic signatures as an alternative to handwritten signatures outside the usual areas.
In this work, we devote our attention to the exploration of the legal preconditions relevant for the EU, technological preconditions, review of the existing software, and a proposal of our own software platform.
The result is an open software platform for operating systems Windows, macOS, and Linux.
Using this application, it is possible, following the eIDAS Regulation, to sign and verify the signature of any document with the legal effect of an attested signature.
Our solution supports ID cards of the Slovak Republic with the eID chip without any modifications or settings, which enables use outside the web applications of the state.
The application is not tied only to the Slovak eID but is ready for further use with various devices in other countries and as a platform for other types of signatures.
There is also a website that, in addition to the possibility of downloading software and brief documentation, provides general information about electronic signatures.
The emphasis is on the simple extensibility of the program via language translations, documentation, functionality, and general information.
The thesis can be further extended with novel ways of signing or by improving flexibility, which is not completely up to par with the compared software.
}

@@

While, specifically in the European Union, electronic signatures seem to be used relatively extensively in communication with the government, they do not seem to have been widely adopted in exchanges between individuals or smaller organizations, which still seem to prefer handwritten signatures.
In many cases, however, they may be impractical for electronic communication, due to the need of a person's physical presence or the printing and sending of documents.

Our goal is to explore an accessible way of using electronic signatures.
In the beginning, we go over the preconditions that make it a viable alternative to handwritten signatures.
We cover the legal preconditions relevant to the European Union, and Slovakia in particular.
Also, we explain the technical preconditions, focusing on cryptographic concepts like asymmetric cryptography, timestamping, hashing, and standards that comply with the laws of the European Union.
Secondly, we do a review of the existing software, where we briefly cover several applications that are meant for basic use.
We compare them using several criteria to find out the strengths and weaknesses of the current implementations and identify a potential gap.
Finally, we devote our attention to our own software's implementation.
We explain the proposed desktop software and go into details on software engineering aspects.
Specifically, we cover usability, localization, modularity, architecture, testing, and automation.
In the end, we compare it to the existing software and identify possible future improvements.
Another result we focus on is a supporting website and how it helps with the use of the software, all in a way that is open for extension by the community.

@@

We did a review of the existing software, focusing on licensing, cross-platform compatibility, localization, file support, ease of setup, flexibility, verification ability, and extensibility.
We identified the positive and negative sides of the current implementations and found a gap in the software for signing using qualified electronic signatures.
